---
title: "A Rebuttal to Watters et al. (2020): Misinterpretation of Predator-Prey Dynamics leading to erroneous Precautionary Management in Antarctic Krill Fisheries"
author:
  - name: "Andrew Lowther"
    affiliation: 1
  - name: "Ulf Lindstrøm"
    affiliation: 2
  - name: "Bjørn Krafft"
    affiliation: 3
  - name: "Javier A. Arata"
    affiliation: 4
    
affiliation:
  - id: 1
    name: "Norwegian Polar Institute, Department of Research, Tromsø, Norway"
  - id: 2
    name: "Institute of Marine Research, Tromsø, Norway"
  - id: 3
    name: "Institute of Marine Research, Bergen, Norway"
  - id: 4
    name: "Association of Responsible Krill harvesting companies, ARK, Margate, TAS, Australia"

geometry: margin=3cm
date: "`r Sys.Date()`"
# bibliography: mylib.bib
# linenumbers: true
# numbersections: true
# csl: elsarticle.cls

output: word_document
always_allow_html: true

  # rticles::elsevier_article:
  #   latex_engine: xelatex  # Optional: Use xelatex for better font handling if needed
  #   keep_tex: true  # Optional: Keep the .tex file for debugging
  #   fig_caption: FALSE
    
header-includes:
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \floatplacement{figure}{H}
  - \newcommand{\pandocbounded}[1]{#1}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
  - \usepackage{threeparttablex}
  - \usepackage[normalem]{ulem}
  - \usepackage{makecell}
  - \usepackage{xcolor}
  - \usepackage{booktabs}
  - \newcommand{\blandscape}{\begin{landscape}}
  - \newcommand{\elandscape}{\end{landscape}}
  - \usepackage{float} \floatplacement{figure}{H} 
  - \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}}     -   - \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
#output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment=NA)
#options(knitr.kable.NA = '\\-')
library(trip)
library(raster)
#library(rgdal)
library(lubridate)
library(tidyverse)
library(sf)
library(raster)
#options(kableExtra.latex.load_packages = FALSE)
library(kableExtra)
library(RefManageR)
library(bibtex)
#library(rbbt)
library(pander)
library(track2KBA)
library(knitr)
library(lmerTest)
library(formatR)
library(tinytex)
library(patchwork)
```

# A. Analysis of Survey Coverage

```{r Survey cover effort, fig.width=7.5, fig.asp=0.6, out.width="80%",fig.align="centre",fig.cap="Figure S1: Survey coverage (total transect length surveyed) by gSSMU and year (up). Frequency distribution of total transect length, including the 10th percentile (dashed line) is shown below.", echo=FALSE}

# load script that produces fig_g1 and fig_g2
suppressMessages(
  suppressWarnings(
    source("./A_Results_Suppl/Figure_S1_Survey_analysis.R")
  )
)

print(fig_g1)

print(fig_g2)

```

Figure S1. Survey coverage (total transect length surveyed) by gSSMU and year (up). Frequency distribution of total transect length, including the 10th percentile (dashed line) is shown below.


# B. Imputed summer LKB values


#Table S1: Estimations of mean and sd of Ln(LKB), as per Equation (2) (Watters et al. 2020).  

```{r Table.01, results='asis', warning=FALSE, echo=FALSE, message=FALSE, comment=NA}

# Load script without printing anything to the document
# invisible(
#   capture.output(
#   source("./A_Results_Suppl/Figure1_S2_S5_Table_S1.R"),
#   type = "output"
#   )
# )
# 
# kable(Table.01, digits = 3)


```



```{r Survey data, fig.width=7.5, fig.asp=0.65, out.width="80%",fig.align="centre", warning=FALSE, echo=FALSE}

#Load script without printing anything to the document
invisible(
  capture.output(
  source("./A_Results_Suppl/Figure1_S2_S5_Table_S1.R"),
  type = "output"
  )
)

# print(FigS2_2)
# 
# print(FigS2_1)

```

Figure S2. Krill survey estimates (LKB) for the study period. Top: gSSMU #2 (SSIW-EI); Bottom:gSSMU #1 (BS). Filled circles: Survey data from Watters et al. (2020); Open circles: mean imputed median LKB-value for -/+ SAM; filled triangles: survey data from Wang et al. (2025).


```{r Survey histogram, fig.width=7.5, fig.asp=0.7, out.width="80%",fig.align="centre", warning=FALSE, echo=FALSE}

print(survey_hist)


```

Figure S3. Histogram of krill acoustic biomass data (ln-transformed) by gSMMU (#1 & #2) and SAM (Neg, Pos). Solid line: Log-Normal distribution fitted to survey data.


```{r Imputed Survey, fig.width=6.5, fig.asp=0.65, out.width="80%",fig.align="centre", warning=FALSE, echo=FALSE}

print(imputed_survey_hist)


```

Figure S4. 

Watters et al. extracted the mean ln(K) from a uniform distribution, ranging from  0.1k to 10k. If applied to the mean-ln(LKB) for gSSMU #1 and -SAM (Table S1), the values range from 1.364-136.43 [ln(LKB) ~ U(0.1k, 10k); k = 13.643].
At the same time, the imputed LKB were truncated, so the minimum LKB = 10,000 tonnes (ln(k) = 9.2103) and maximum LKB = 100,000,000 tonnes (ln(lkb) = 18.4207.  This means that imputed mean-ln(LKB) will be larger than the maximum, thus, 1x10^8 tonnes. Same goes for sd(lkb).
As result, the distribution of the imputed ln(LKB) values are skewed up (Figure S4).

```{r Imputed mean-log(LKB), fig.width=6.5, fig.asp=0.65, out.width="80%",fig.align="centre", warning=FALSE, echo=FALSE}

print(imputed_survey_hist.mean)


```

Figure S5. 

The removal of the equation (2), K ~ U(0.1k, 10k), corrects the issue observed in figure S4. 

  
# Citations
