---
title: "A Rebuttal to Watters et al. (2020): Misinterpretation of Predator-Prey Dynamics leading to erroneous Precautionary Management in Antarctic Krill Fisheries"
author:
  - name: "Andrew Lowther"
    affiliation: 1
  - name: "Ulf Lindstrøm"
    affiliation: 2
  - name: "Bjørn Krafft"
    affiliation: 3
  - name: "Javier A. Arata"
    affiliation: 4
    
affiliation:
  - id: 1
    name: "Norwegian Polar Institute, Department of Research, Tromsø, Norway"
  - id: 2
    name: "Institute of Marine Research, Tromsø, Norway"
  - id: 3
    name: "Institute of Marine Research, Bergen, Norway"
  - id: 4
    name: "Association of Responsible Krill harvesting companies, ARK, Margate, TAS, Australia"

geometry: margin=3cm
date: "`r Sys.Date()`"
# bibliography: mylib.bib
# linenumbers: true
# numbersections: true
# csl: elsarticle.cls

output: word_document
always_allow_html: true

  # rticles::elsevier_article:
  #   latex_engine: xelatex  # Optional: Use xelatex for better font handling if needed
  #   keep_tex: true  # Optional: Keep the .tex file for debugging
  #   fig_caption: FALSE
    
header-includes:
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \floatplacement{figure}{H}
  - \newcommand{\pandocbounded}[1]{#1}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
  - \usepackage{threeparttablex}
  - \usepackage[normalem]{ulem}
  - \usepackage{makecell}
  - \usepackage{xcolor}
  - \usepackage{booktabs}
  - \newcommand{\blandscape}{\begin{landscape}}
  - \newcommand{\elandscape}{\end{landscape}}
  - \usepackage{float} \floatplacement{figure}{H} 
  - \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}}     -   - \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
#output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment=NA)
#options(knitr.kable.NA = '\\-')
library(trip)
library(raster)
#library(rgdal)
library(lubridate)
library(tidyverse)
library(sf)
library(raster)
#options(kableExtra.latex.load_packages = FALSE)
library(kableExtra)
library(RefManageR)
library(bibtex)
#library(rbbt)
library(pander)
library(track2KBA)
library(knitr)
library(lmerTest)
library(formatR)
library(tinytex)
library(patchwork)
library(tibble)
library(ggplot2)
library(scales)
library(stringr)
```

# A. Analysis of Survey Coverage

Survey coverage, expressed as nautical miles surveyed within each gSSMU
and survey leg, varied among years and between legs. In most years, two
survey legs were conducted. Watters et al. (2020) proposed excluding
surveys with transect coverage of 80 nm or less, but did not apply this
criterion.

Rather than applying a fixed threshold, we examined the frequency
distribution of surveyed transect lengths for gSSMUs 1 and 2 separately
and excluded surveys falling below the lower 10th percentile, rounded up
to the nearest 10 nm. This approach resulted in the exclusion of three
survey legs in gSSMU 1 (≤10 nm; 1996, 1999, 2010, and 2012) and four
survey legs in gSSMU 2 (≤200 nm; 2010, 2011, 2012, and 2015) (Figure
S1).

```{r Survey cover effort, fig.width=7.5, fig.asp=0.6, out.width="80%",fig.align="centre",fig.cap="Figure S1: Survey coverage (total transect length surveyed) by gSSMU and year (up). Frequency distribution of total transect length, including the 10th percentile (dashed line) is shown below.", echo=FALSE}

# load script that produces fig_g1 and fig_g2
suppressMessages(
  suppressWarnings(
    source("./A_Results_Suppl/Figure_S1_Survey_analysis.R")
  )
)

print(fig_g1)

print(fig_g2)

```

Figure S1. Annual survey coverage (total transect length surveyed) by
gSSMU, year and survey leg [top panel]. Frequency distribution of summed
transect lengths across all surveys, with the 10th percentile marked by
the dashed line [bottom panel].

\newpage
# B. Distribution of observed vs imputed summer ln*LKB* values

We evaluated the procedure used to impute LKB values in years without
survey data. First, we assessed whether ln*LKB* from survey observations
followed an appropriate parametric distribution. We then generated
imputed ln*LKB* values following equations (1) and (2) of Watters et al.
(2020) and compared the resulting distributions with the empirical
distribution of observed data.

We fitted a log-normal distribution to the survey-based LKB data and
assessed normality of ln*LKB* using the Shapiro--Wilk test. Using the
same parameterization, we generated 100,000 imputed ln*LKB* values from
equations (1) and (2), applying lower and upper bounds of 10,000 t and 1
× 10⁹ t, respectively, as defined in Supplementary Material S1.
Normality of the imputed ln*LKB* distribution was then re-evaluated
using the Shapiro--Wilk test.

Observed survey-based LKB distributions for all gSSMU--SAM combinations
were well described by a log-normal distribution (Supplementary
Material, Table S1; Figure S2). In contrast, ln(LKB) values generated
through the imputation procedure of Watters et al. (2020) were
concentrated toward the upper bound of the imposed limits, with mean
ln(LKB) values of 17.5--17.7 (approximately 44 Mt), and did not conform
to a log-normal distribution (Supplementary Material, Table S2; Figure
S3).

Table S1. Shapiro--Wilk test results for normality of observed
log-transformed LKB values, reported by gSSMU and SAM sign. Columns show
sample size (n), p-value, and W-statistic.

```{r Table S1, results='asis', warning=FALSE, echo=FALSE, message=FALSE, comment=NA}

# load script that produces fig_g1 and fig_g2
suppressMessages(
  suppressWarnings(
    source("./A_Results_Suppl/Figure_S2_S4_Table_S1_S2.R")
  )
)

kable(shapiro_results1, digits = 3)

```

Table S2. Shapiro--Wilk test results for normality of imputed
log-transformed LKB values. The table reports the total number of
imputations generated (n_total), the subsample used for the test
(n_used), and the corresponding W and p values for each gSSMU and SAM
sign.

```{r Table S2, results='asis', warning=FALSE, echo=FALSE, message=FALSE, comment=NA}

kable(shapiro_imp, digits = 3)

```

```{r Survey histogram, fig.width=7.5, fig.asp=0.7, out.width="80%",fig.align="centre", warning=FALSE, echo=FALSE}

# load script that produces fig_g1 and fig_g2
suppressMessages(
  suppressWarnings(
    source("./A_Results_Suppl/Figure_S2_S4_Table_S1_S2.R")
  )
)
print(survey_hist)


```

Figure S2. Distribution of ln-transformed acoustic biomass (LnSurvey)
shown as histograms stratified by gSSMU (1, 2) and SAM sign (Neg, Pos).
The solid curve in each panel shows the log-normal density fitted to the
observed survey values.

```{r Imputed Survey, fig.width=6.5, fig.asp=0.65, out.width="80%",fig.align="centre", warning=FALSE, echo=FALSE}

print(imputed_survey_hist)


```

Figure S3. Distribution of ln-transformed imputed biomass (LnSurvey)
shown as histograms stratified by gSSMU (1, 2) and SAM sign (Neg, Pos).
The solid curve in each panel shows the log-normal density fitted to the
observed survey values.

\newpage
# Citations

Watters, George M., Jefferson T. Hinke, and Christian S. Reiss. 2020.
\"Long-Term Observations from Antarctica Demonstrate That Mismatched
Scales of Fisheries Management and Predator-Prey Interaction Lead to
Erroneous Conclusions about Precaution.\" Scientific Reports 10 (1).
<https://doi.org/10.1038/s41598-020-59223-9>.
